---
title: "Lab 9"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(stringr)
library(tidyverse)
```
# Basic matches

```{r}
x <- c("apple", "banana", "pear")
str_view(x, "an")
```

```{r}
str_view(x, ".a.")
```

```{r}
# To create the regular expression, we need \\
dot <- "\\."

# But the expression itself only contains one:
writeLines(dot)
```

```{r}
# And this tells R to look for an explicit .
x <- c("abc", "a.c", "bef")
str_view(x, "a\\.c")
```
```{r}
x <- "a\\b"
writeLines(x)
```

```{r}
str_view(x, "\\\\")
```

# 14.3.1.1 Exercises 

#Explain why each of these strings don’t match a \: "\", "\\", "\\\".
# \ is read as an escape, so you have to escape the \ before it will be read as an actual \. This would create the expression \\, but this also needs a string which needs an escape. So the \\\ is making a string and escaping it from itself, so that the fourth \ is read as a \

# The ^ symbol will match with the beginning of the string, and the $ symbol will match the end of a string

```{r}
x <- c("apple", "banana", "pear")
str_view(x, "^a")
```

```{r}
str_view(x, "a$")
```
 


```{r}
x <- c("apple pie", "apple", "apple cake")
str_view(x, "apple")
```

```{r}
str_view(x, "^apple$")
```

# 14.3.2.1 Exercises
#How would you match the literal string "$^$"?
# You would match the string $^$ as something that starts and ends with the same letter


### Special Patterns
# \d : matches any digit.
# \s : matches any whitespace (e.g. space, tab, newline).
# [abc] : matches a, b, or c.
# [^abc] : matches anything except a, b, or c.
# Remember, to create a regular expression containing \d or \s, you’ll need to escape the \ for the string, so you’ll type "\\d" or "\\s"


```{r}
x <- c("abc", "a.c", "a*c", "a c")
str_view(x, "a[.]c")
```

```{r}
str_view(x, ".[*]c")
```

```{r}
str_view(x, "a[ ]")
```

# | gives you either side of it but not both mixed together. abc | xyz will match abc or xyz, but not abxyz
```{r}
x <- c("grey", "gray")
str_view(x, "gr(e|a)y")
```

### 14.3.3.1 Exercises
# Create regular expressions to find all words that:
# Start with a vowel.
# That only contain consonants. (Hint: thinking about matching “not”-vowels.)
# End with ed, but not with eed.
# End with ing or ise.
```{r}
x <- c("orange", "cat", "table", "airplane", "water", "igloo", "orangutan", "earring", "umbrella", "octopus", "window", "ambulance")
str_view(x, "^a") 
str_view(x, "^e")
str_view(x, "^i")
str_view(x, "^o")
str_view(x, "^u")

```
```{r}
x <- "1888 is the longest year in Roman numerals: MDCCCLXXXVIII"
str_view(x, "CC?")
```
```{r}
str_view(x, "CC+")
```
```{r}
str_view(x, 'C[LX]+')
```
 
```{r}
str_view(x, "C{2}")
```

```{r}
str_view(x, "C{2,}")
```


```{r}
str_view(x, 'C{2,3}?')
```

```{r}
str_view(x, 'C[LX]+?')
```

# Grouping and backreferences
```{r}
str_view(fruit, "(..)\\1", match = TRUE)
```

# Detect matches

```{r}
x <- c("apple", "banana", "pear")
str_detect(x, "e")
```

```{r}
# How many common words start with t?
sum(str_detect(words, "^t"))
```

```{r}
# What proportion of common words end with a vowel?
mean(str_detect(words, "[aeiou]$"))
```

```{r}
# Find all words containing at least one vowel, and negate
no_vowels_1 <- !str_detect(words, "[aeiou]")
```

```{r}
# Find all words consisting only of consonants (non-vowels)
no_vowels_2 <- str_detect(words, "^[^aeiou]+$")
identical(no_vowels_1, no_vowels_2)
```

```{r}
words[str_detect(words, "x$")]
```
```{r}
str_subset(words, "x$")
```

```{r}
df <- tibble(
  word = words, 
  i = seq_along(word)
)
df %>% 
  filter(str_detect(word, "x$"))
```

```{r}
x <- c("apple", "banana", "pear")
str_count(x, "a")
```

```{r}
# On average, how many vowels per word?
mean(str_count(words, "[aeiou]"))
```

```{r}
df %>% 
  mutate(
    vowels = str_count(word, "[aeiou]"),
    consonants = str_count(word, "[^aeiou]")
  )
```

```{r}
str_count("abababa", "aba")
```

```{r}
str_view_all("abababa", "aba")
```

# Extract matches

