---
title: "Lab 3"
author: "Brooke Linnehan"
date: "9/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown


```{r}
library(tidyverse)

```


```{r}
download.file(url="https://ndownloader.figshare.com/files/2292169",

destfile = "data/portal_data_joined.csv") 
```



```{r}
surveys <- read_csv("data/portal_data_joined.csv")
```

### Size
```{r}
dim(surveys)
nrow(surveys)
ncol(surveys)

```
### Content
```{r}
head(surveys)
tail(surveys)
```

### Names
```{r, eval = FALSE}
names(surveys)
rownames(surveys)
```


### Summary
```{r, eval=FALSE}
str(surveys)

summary(surveys)
```

### Challenge 1: Based on the output of str(surveys):

# What is the class of the object surveys? 
# A: data frame
# How many rows and how many columns are in this object?
# A: 34,786 rows; 13 columns
# How many species have been recorded during these surveys?
# A: 48 species


### Challenge 2

# 1. Create a data.frame (surveys_200) containing only the data in row 200 of the surveys dataset.
```{r}
surveys_200<-surveys[200, ]
data.frame(surveys_200)
n_rows <- nrow(surveys_200)
n_rows

```

```{r}
tail(surveys)
```

# 2

```{r}
n_rows <- nrow(surveys)
surveys_last <- surveys[n_rows, ]
surveys_last
```
# 3

```{r}
surveys_middle <- surveys[n_rows / 2, ]
surveys_middle
```

# 4
```{r}
surveys_head <- head(surveys[-(7:n_rows), ])
surveys_head
```

### Factors


```{r}
sex<- factor(c("male", "female"))


levels(sex)
nlevels(sex)

```
# Current Order
```{r}
sex

```
# After Reordering
```{r}
sex <- factor(sex, levels = c("male", "female"))
sex
```

### Converting Factors
```{r}
as.character(sex)

```

```{r}
year_fct <- factor(c(1990, 1983, 1977, 1998, 1990))
as.numeric(as.character(year_fct))
as.numeric(levels(year_fct))[year_fct]
```
### Renaming Factors
```{r}
plot(as.factor(surveys$sex))

```

```{r}
sex<- factor(surveys$sex)
head(sex)
levels(sex)
levels(sex)[1]<- "undetermined"
levels(sex)
head(sex)
```


### Challenge
```{r}
levels(sex)[2:3] <- c("female", "male")
sex <- factor(sex, levels = c("female", "male", "undetermined"))
plot(sex)
```



## Compare the difference beween our data read as "factor" vs "character"
```{r}
surveys <- read.csv("data/portal_data_joined.csv", stringsAsFactors = TRUE)
str(surveys)
surveys <- read.csv("data/portal_data_joined.csv", stringsAsFactors = FALSE)

str(surveys)
## Convert the column "plot_type" into a factor
surveys$plot_type <- factor(surveys$plot_type)

```



# Challenge
## 1
```{r}
animal_data <- data.frame(
  animal = c("dog", "cat", "sea_cucumber", "sea_urchin"),
  feel = c("furry", "soft", "squishy", "spiny"),
  weight = c(45, 8, 1.1, 0.8)
)
```
## 2

### class: country and climate are character, temperature is numeric, and northern_hemisphere and has_kangaroo are logical
```{r}
country_climate <- data.frame(
  country = c("Canada", "Panama", "South_Africa", "Australia"),
  climate = c("cold", "hot", "temperate", "hot/temperate"),
  temperature = c(10, 30, 18, 15),
  northern_hemisphere = c(TRUE, TRUE, FALSE, FALSE),
  has_kangaroo = c(FALSE, FALSE, FALSE, TRUE)
)
```
```{r}
str(country_climate)
```

### The classes are not what I expected, because temperature should be numeric, and northern_hemisphere and has_kangaroo should be logical. 
### If I added stringsasfactors = FALSE, it would have made character vectors and not factors
### Temperature and nothern_hemisphere should not have quotes, and has_kangaroo should say TRUE instead of 1


# Formatting Dates

###Using "Lubridate" Package from tidyverse
```{r}
library("lubridate")
```

```{r}
my_date <- ymd("2015-01-01")
str(my_date)

```

```{r}
my_date <- ymd(paste("2015", "1", "1", sep = "-"))
str(my_date)

```
### Applying faunction to surveys dataset

```{r, eval = FALSE}
paste(surveys$year, surveys$month, surveys$day, sep = "-")

```
```{r, eval = FALSE}
ymd(paste(surveys$year, surveys$month, surveys$day, sep = "-"))
```
```{r, eval = FALSE}
surveys$date <- ymd(paste(surveys$year, surveys$month, surveys$day, sep = "-"))
str(surveys)
summary(surveys$date)
missing_dates <- surveys[is.na(surveys$date), c("year", "month", "day")]
head(missing_dates)
```


