---
title: "Lab 5"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(lubridate)
```

```{r}
Confirmed_State_3_13 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-13-2020.csv")) %>%
  rename(Country_Region = "Country/Region", Province_State = "Province/State") %>% 
  filter (Country_Region == "US") %>% 
  group_by(Province_State, Country_Region) %>% 
  summarise(Confirmed = sum(Confirmed)) 
str(Confirmed_State_3_13)
```

```{r}
Confirmed_State_9_13 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-13-2020.csv")) %>% 
  filter (Country_Region == "US") %>% 
  group_by(Province_State, Country_Region) %>% 
  summarise(Confirmed = sum(Confirmed)) 
str(Confirmed_State_9_13)
Confirmed_State_9_13 <- Confirmed_State_9_13 %>% 
  filter(Province_State != "Recovered") 
```


```{r}
setdiff(Confirmed_State_9_13$Province_State, Confirmed_State_3_13$Province_State)

```

# Full Join to Merge Tables
```{r}
Confirmed_State_3_13_9_13_joined <- full_join(Confirmed_State_3_13,
      Confirmed_State_9_13, by = c("Province_State"))
head(Confirmed_State_3_13_9_13_joined)
tail(Confirmed_State_3_13_9_13_joined, 5)
which(is.na(Confirmed_State_3_13_9_13_joined))
```

```{r}
Confirmed_State_3_13_9_13_joined <- full_join(Confirmed_State_3_13,
      Confirmed_State_9_13, by = c("Province_State")) %>% 
      rename(Confirmed_3_13_2020 = "Confirmed.x", Confirmed_9_13_2020 = "Confirmed.y") %>% 
      select(-Country_Region.x, -Country_Region.y) %>% 
      replace_na(list(Confirmed_3_13_2020 = 0))
head(Confirmed_State_3_13_9_13_joined)
which(is.na(Confirmed_State_3_13_9_13_joined))
```

### Switching between wide and long table formats

```{r}
Confirmed_State_3_13_9_13_joined_long <- Confirmed_State_3_13_9_13_joined %>% 
              pivot_longer(-c(Province_State),
                            names_to = "Date", values_to = "Confirmed")
ggplot(Confirmed_State_3_13_9_13_joined_long, aes(x = Confirmed,  y = Province_State))  + 
    geom_point(aes(color = Date)) 
```

# Wide Format

```{r}
time_series_confirmed <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")) %>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")
head(time_series_confirmed)
```

# Converting to Long Format

```{r}
time_series_confirmed_long <- time_series_confirmed %>% 
               pivot_longer(-c(Province_State, Country_Region, Lat, Long),
                            names_to = "Date", values_to = "Confirmed") 
head(time_series_confirmed_long)
```

# Creating Another Table with

```{r}
download.file(url="https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv", 
               destfile = "data/time_series_covid19_deaths_global.csv")

```

```{r}
time_series_deaths <- read_csv("data/time_series_covid19_deaths_global.csv")%>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")

```

```{r}
time_series_deaths_long <- time_series_deaths %>% 
               pivot_longer(-c(Province_State, Country_Region, Lat, Long),
                            names_to = "Date", values_to = "Deaths") 
head(time_series_deaths_long)
```


```{r}
time_series_confirmed_long <- time_series_confirmed_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".", remove = FALSE)
head(time_series_confirmed_long)
```
# Creating Key

```{r}
time_series_deaths_long <- time_series_deaths_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".") %>% 
  select(Key, Deaths)
```

# Joined Data for Confirmed and Deaths, Removing Key

```{r}
time_series_long_joined <- full_join(time_series_confirmed_long,
              time_series_deaths_long, by = c("Key")) %>% 
              select(-Key)
head(time_series_long_joined)

```
```{r}
which(is.na(time_series_long_joined$Confirmed))
which(is.na(time_series_long_joined$Deaths))
```

# Reformat Dates

```{r}
time_series_long_joined$Date <- mdy(time_series_long_joined$Date)
```

# Creating a Column to Graph Both Cofirmed and Deaths
```{r}
time_series_long_joined_counts <- time_series_long_joined %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long, Date),
               names_to = "Report_Type", values_to = "Counts")
head(time_series_long_joined_counts)
```

# Making Graphs From Time Series Data

```{r}
time_series_long_joined %>% 
  group_by(Country_Region,Date) %>% 
  summarise_at(c("Confirmed", "Deaths"), sum) %>% 
  filter (Country_Region == "US") %>% 
    ggplot(aes(x = Date,  y = Deaths)) + 
    geom_point() +
    geom_line() +
    ggtitle("US COVID-19 Deaths")
```

```{r}
time_series_long_joined %>% 
  group_by(Country_Region,Date) %>% 
  summarise_at(c("Confirmed", "Deaths"), sum) %>% 
  filter (Country_Region %in% c("China","Japan", "Korea, South",
                                "Italy","Spain", "US")) %>% 
    ggplot(aes(x = Date,  y = Deaths)) + 
    geom_point() +
    geom_line() +
    ggtitle("COVID-19 Deaths") +
    facet_wrap(~Country_Region, ncol=2, scales="free_y")
```

```{r}
time_series_long_joined %>% 
    group_by(Country_Region,Date) %>% 
    summarise_at(c("Confirmed", "Deaths"), sum) %>% 
    filter (Country_Region %in% c("China","France","Italy", 
                                "Korea, South", "US")) %>% 
    ggplot(aes(x = Date,  y = Deaths, color = Country_Region)) + 
    geom_point() +
    geom_line() +
    ggtitle("COVID-19 Deaths")
```

```{r}
time_series_long_joined_counts %>% 
  group_by(Country_Region, Report_Type, Date) %>% 
  summarise(Counts = sum(Counts)) %>% 
  filter (Country_Region == "US") %>% 
    ggplot(aes(x = Date,  y = log2(Counts), fill = Report_Type, color = Report_Type)) + 
    geom_point() +
    geom_line() +
    ggtitle("US COVID-19 Cases")
```


### Exercises


### Exercise 1 in "Data Visualization" file


### Exercise 2


### Exercise 3


### Exercise 4 Plot the total number of confirmed deaths per day worldwide

```{r, eval = FALSE}
time_series_deaths_long %>% 
group_by(Country_Region) %>% 
  summarise_at(c("Deaths"), sum) %>% 
  ggplot(aes(x = Date, y = Deaths)) +
    geom_line() +
    ggtitle("Covid-19 Deaths Worldwide")
```
```{r}
time_series_deaths_long
```


### Exercise 5

```{r}

```

### Exercise 6 Plot US deaths/confirmed cases per day.
```{r}
time_series_long_joined_counts %>% 
  group_by(Country_Region, Report_Type, Date) %>% 
  summarise(Counts = sum(Counts)) %>% 
  filter (Country_Region == "US") %>% 
    ggplot(aes(x = Date,  y = log2(Counts), fill = Report_Type, color = Report_Type)) + 
    geom_point() +
    geom_line() +
    ggtitle("US COVID-19 Cases")
```

### Exercise 7 Make a single graph with from the times series data with the 10 countries with the hightest death total.

```{r}
top10c <- time_series_long_joined %>% 
  group_by(Country_Region, sum(Deaths), Date) %>% 
  summarize(Confirmed = sum(Confirmed), Deaths = sum(Deaths), min_Deaths = min(Deaths)) %>% 
  arrange(desc(Country_Region)) %>% 
  slice(1:10)

ggplot(data = top10c, aes(x = Date, y = Deaths, color = Country_Region)) +
  geom_point() +
geom_line() +
  ggtitle("COVID-19 Deaths Top 10 Countries")

  

```



